---
title: "Tutorial"
author: "Wendy Leuenberger"
date: "`r Sys.Date()`"
output: 
  html_document:
    keep_md: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```


# Tutorial with the repository

This tutorial is designed for step-by-step editing of a GitHub repository and for learning how to submit jobs to the HPCC. The repository is meant to pick up at the end of HPCC.Rmd/md/html, once you have your computer configured to allow GitHub and RStudio to talk to each other and have the HPCC account mapped to your drive. You can also work through the exercises without the HPCC. 

## Put this repository on your computer

Fork the repository: Only owners of a repository can make changes. But if the repository is public, others can "fork" the repository. This creates a local copy on their GitHub that they can make changes to. Go to the main page of this repository. Above the green `<> Code` button, there are four buttons. The third says `Fork`. Click this button and make a copy of this repository in your account. 

## Clone the repository to your computer

Create a new RStudio project with Version Control and use your forked repository's url to link the project to your repository. You can find the url by clicking the green  `<> Code` button. (See the HPCC files in the Tutorial folder for more detail). Leave this project open. 

## Learn how to make edits

There are multiple ways to make edits to a GitHub repository. Try out these three ways:

1. Commit using the web browser. 

    - In your web browser, navigate to the main page of the repository. Your README file should be showing.
    
    - Above the header, you can see README on the left and a pencil and outline buttons on the right. Click on the pencil to edit. Alternatively, you can click the file name and then the edit button on the right. 

    - Replace my name with your name.

    - Click the green `Commit changes...` button on the upper right.
  
    - Add a commit message that is useful - i.e. "Added my name"

    - Leave the toggle on `Commit directly to the main branch`

    - Click the green `Commit changes` button. 

    - Your name will now appear on the main README page!
    
    - Go to your RStudio window that is linked to your GitHub repository. Go to the Terminal window (usually the tab next to the Console in the lower left window). You'll see the file path to your repository. Type `git pull` to pull the change you just made to the README.md in the browser. 

2. In RStudio, using the Git tab. 

    - Open the README.md document. You can find it under Files in the lower right window. 

    - Underneath your name, replace `Today's Date` with the Date. 
    
    - Save the file. 
    
    - In the upper right window, go to the Git tab. Check the box under the `Staged` column for the README.md. 
    
    - Click the `Commit` button above the files.
    
    - Type a commit message like `Added today's date` in the window on the upper right. 
    
    - Click `Commit`
    
    - On the upper right, click the `Push` button.
    
    - Close the Git window and go to the repository in your web browser. Refresh and the date that you added will show up. 
    
3. In RStudio, using command line. 

    - Go back to your README.md in RStudio. 
    
    - Replace my email with your email. 
    
    - Go to the Terminal tab next to the console. (On windows, you can use Ctrl + Alt + M to navigate to the terminal)
    
    - Type `git pull` (you haven't done anything that needs to be pulled, but it's a good practice to pull before pushing in case of changes)
    
    - Type `git add README.md`
    
    - Type `git commit -m 'Added favorite animal'`
    
    - Type `git push`
    
    - Typing `git status` is also helpful for seeing what files have been changed and if you're ahead of or behind the uploaded Git repository. 

    - RStudio is pretty cool in that it allows R and/or other languages in chunks. Here are the bash commands for committing changes. In windows, you can press `Ctrl + Alt + Enter` to run each line of code in the terminal. 
    
```{bash}
git pull

git add README.md

git commit -m 'Added email'

git push

git status
```

    
All of these methods work! The rest of the tutorial will be in command line, but most could be done with any method. Command line is particularly useful with the HPCC, as you can have multiple terminal windows for your computer, the HPCC, and the mapped drive, all within one RStudio window.

## Log onto the HPCC

If you have an HPCC account, open up a second terminal window. 

- Click on Terminal 1 under the Terminal tab

- Select New Terminal (or use `Alt + Shift + R` on Windows)

- Log onto the HPCC by typing `ssh -XY YourNetID@hpcc.msu.edu`. Note the hpcc.msu.edu instead of your regular email, and replace YourNetID with your own net ID. For example, I type: 
    
```{bash}
ssh -XY leuenbe9@hpcc.msu.edu
```
 
- Type your MSU password. It's the same as your email. You won't be able to see the letters as you type. You can also put your password in the bash chunk and press `Ctrl + Alt + Enter`, but be careful to delete the password afterwards/keep it secure. 

- Go to one of the development nodes. I use either `dev-intel16` or `dev-intel18`. Go for whichever has lower usage in parentheses. 

```{bash}
ssh dev-intel16
ssh dev-intel18
```

## Navigate to your mapped drive

- Open another terminal. 

- Change directories to your mapped drive. You may need to go to it in your finder window (file explorer) and enter your password if it doesn't let you. Note which letter you mapped your drive to. Mine is on the `Y` drive

```{bash}
cd /y
```

## Clone the repository on your mapped drive. 

- Get the URL from the green `<> Code` chunk on the main page of the repository in your browser.

- Change the following code to match your URL. In the Terminal window that is linked to your mapped drive, run the following code. 

```{bash}
git clone https://github.com/wleuenberger/GitHub_HPCC_Tutorial.git
```

- Type `ls` and note that your repository is now present on your mapped drive. Change your directory to the repository with `cd GitHub_HPCC_Tutorial`

```{bash}
ls
cd GitHub_HPCC_Tutorial
```

- Switch tabs to the tab where you're logged into the HPCC. You can click the arrows or Terminal # drop down on the upper right of the tab. Or on Windows, you can press `Shift + Alt + F11` or `Shift + Alt + F12` to move between terminal tabs that you have open. 

- Type `ls` to see that the tutorial is already on the HPCC (because you put it on your mapped drive). Change your directory to this folder in this tab as well.

```{bash}
ls
cd GitHub_HPCC_Tutorial
```

## Working with R via the HPCC. 

There are other files in this repository. I've included some data and code to get us started. There are four ways I know of to work with R code on the HPCC:

- OnDemand - open up an RStudio session in your web browser. Lowest barrier to entry, but limited on what you can do. We won't go over this method.

- Open up R in your HPCC terminal and run code directly.

- Run an Rscript on the development node.

- Submit jobs to the HPCC to run remotely

    - This is the one that really makes using the HPCC worthwhile, as you can have multiple jobs running simultaneously for up to a week
    
### Open R in the terminal

You can open R directly and run code line by line similar to RStudio. You have to load certain modules first. This stage is particularly useful for checking to make sure that the code works on the HPCC and that all packages are loaded. 

Load modules
```{bash}
module load R-bundle-CRAN/2023.12-foss-2023a
```

Open an R session on the HPCC
```{bash}
R
```

Now you can type code or use `Ctrl + Alt + Enter` to run code from a file open in RStudio. Try it out for a bit. 

See if you can load some of your common packages, and if not, install them here. If you install them here, they will be available when you submit jobs. 

```{R}
library(tidyverse)
# install.packages('tidyverse')  # I think this one will be loaded, but not all will. 
```

To quit the R session, use the quit function `q()`
```{r}
q()
n # Use n to not save the workspace image
```

### Run a script on command line
```{bash}
Rscript Code/WarblerForHPCC.R
```

It should have printed results. That's great, but what do you do with those results? Note that just running code doesn't inherently mean that anything has been saved for you to work with once the job has completed. 

## Using GitHub and HPCC together

### Modify the R code to save model object

In RStudio, add this code to the bottom of the `WarblerForHPCC.R` file

```{r}
save(out, file = file.path(getwd(), 'Output/out.RData'))
```

Now, save your `WarblerForHPCC.R` file. Go to your terminal tab that's linked to your computer's copy of the GitHub repository. Pull from GitHub to start for safety, then add, commit (include a message), and push the change to GitHub. Switch to the terminal tab that's linked to your mapped drive of the HPCC, and pull the changes there. 

Run the Rscript command again and check that it created the out.RData in the Output folder.

```{bash}
Rscript Code/WarblerForHPCC.R
ls Output/
```

### Add files from the HPCC to GitHub

You can move files both directions (Computer -> GitHub -> HPCC; HPCC -> GitHub -> Computer), assuming they're not too big for GitHub (<100 MB). Add out.RData to GitHub from the HPCC's terminal window, commit and push it, and pull it onto your computer. 

Load it into your R session to see that it's there
```{r}
load(file = 'Output/out.RData')
summary(out)
```

### Bonus: Dynamic saving

It's easy to overwrite files on the HPCC. You won't get a warning or anything. If you anticipate running code multiple times to save objects with the same name, but not overwriting, then it helps to have code that can change the name of the output object based on the number of objects already in the file. Here's some sample code that can help.

This step might be particularly useful if you are running multiple chains of a Bayesian model in parallel, or just want to keep old versions of something while trying something new.

Add this to the R code and put everything on GitHub and the HPCC. (You can skip this step if you don't think you'll use it).

```{r}
ID <- paste0('out',
              length(list.files(path = file.path(getwd(), 'Output/'),
                                pattern = 'out',
                                full.names = FALSE)) + 1)

# Print output file name in slurm.out file
# Helpful for keeping track if you're running multiple versions
paste('\n ************************************** \n \n',
      'Output File Name:', ID, '\n \n',
      '**************************************
      ') %>% cat

save(out, file = file.path(getwd(), 'Output', paste0(ID, '.RData')))

```

## Submit a job to the HPCC

Take a look at runWarbler.sb in the main section of the repository. This is the file that tells the HPCC to run a job. You can open it in RStudio or in Notepad++. You can set the time, memory, and more. Change the email address on line 20 to your email address so that you are emailed when the job finishes or if it fails. 

Generally, I change the time and memory pretty often and the rest largely stays the same. Make sure to update the line with your code file if you want to run different code. The file path is relative to the location of the sbatch file. I usually leave the sbatch file in the main part of my repository. 

Note: Please change the email address in line 20 to your email instead of mine. 

Submit the job to the HPCC using sbatch and the file name. 
```{bash}
sbatch runWarbler.sb
```

Take a look at your queue of submitted jobs. This model runs quickly so it might have already finished. If so, you can run it again and immediately run this line. 
```{bash}
sq
```

```{bash}
sbatch runWarbler.sb
sq
```

The number that it tells you is the job ID. This number can help you keep track of your jobs and look at the output file that the job produces. Take a look here, but change the number to match your job number
```{bash}
cat slurm-33085536.out
```

You can see the model output and the saved model name here. Take a look in the repo.
```{bash}
ls Output/
```

### HPCC Bits and Pieces

Sometimes you need to cancel a job. You can run scancel with your job number
```{bash}
scancel 33085536
```

Take a look at your quota of file space. Check the ICER website if you need to add more space
```{bash}
quota
```

Sometimes it's hard to know how much time and memory a job will take. You can look at past jobs to adjust future runs with reportseff
```{bash}
reportseff -u YourNetID
reportseff -u leuenbe9
```

